---
title: 'Hash\: dictionaries for R'
author: "Liz Sander, Civis Analytics"
date: "August 23, 2017"
fontsize: 10pt
output:
  beamer_presentation:
    colortheme: beaver
    fonttheme: structurebold
    incremental: yes
    theme: Boadilla
  ioslides_presentation:
    incremental: yes
    logo: CIVIS_ICON_0116.png
  slidy_presentation:
    incremental: yes
    logo: CIVIS_ICON_0116.png
---

## What is a dictionary?
- Also known as a hashmap
- Set of key:value pairs

## What is a dictionary?
```{r}
library(hash)
mydict <- hash('a'=1:3,
               'b'=letters[1:3],
               'c'=matrix(1:8,2,4))
print(mydict)
```

## Okay, so it's a named list?
```{r}
mylist <- list('a'=1:3,
               'b'=letters[1:3],
               'c'=matrix(1:8,2,4))
mylist[['a']]
mydict[['a']]
```

## Okay, so it's a named list?
```{r}
mylist$a
mydict$a
mylist[[1]]
# mydict[[1]]
```

## Okay, so it's a named list?
- Dictionaries aren't ordered. You can only access values by their keys
- Trades known order for speed and convenience

# Why dictionaries?

## Easy to look things up
- Doesn't matter for the user
- Faster for the computer
- With a list, the computer has to scan through the list in order
    - (up to length(mylist) operations)
    - Like flipping page by page through a (physical) dictionary until you 
    find the word you want
- With a dict, the computer runs an algorithm on the key to find 
  where it put the value (only one operation)
    - Like thinking for a second about where in the alphabet your word is, then
    flipping to the correct page right away

## Easy to look things up
- This key:value mapping algorithm is called hashing!
- Fun fact: Have you heard of hashing for passwords or cryptography? Same thing! Hashing is awesome :)
- This computer memory stuff is nitty-gritty, but these details are what make
dicts useful.

## Easy to add things
- Doesn't matter for the user
```{r, eval=FALSE}
mydict['d'] = 100
mylist['d'] = 100
```

## Easy to add things
- Way easier for the computer!
- When you add something to the list, the computer has to:
    - make a new, slightly bigger list
    - copy the old list over
    - tack your new thing on at the end
- The computer only makes as much space for the list as it needs right then
    - Makes a perfect size box, and fills it up all of the way.
- A dictionary is more like a cabinet with lots of drawers, some of them empty
    - Adding a new thing just means finding an empty drawer

## Easy to remove things
- Easy for the user (although not identical)
```{r}
mylist = mylist[names(mylist) != 'b']
print(names(mylist))
del('b', mydict) # no need to reassign
print(names(mydict))
```

## Easy to remove things
- With lists, you may have to scan through the whole list/create a whole new list, depending on method
- Dicts just take the item out of the drawer and leave the drawer empty

## When should I use dicts?
Lists are great for\:

- using lapply()
- keeping things in a specific order

Dicts are great for\:

- Speedy access, addition, and deletion of elements
- Variables you use and change a lot
- Data where order doesn't matter

## When should I use dicts?
- keeping track of counts (e.g., counting occurrences of words in text)
- lookup table to store results from slow calculations, so you don't have to calculate more than once
- probably most places you'd otherwise use a list
- nested dictionaries are a thing (think list of lists, but faster)

## Code Example!
